New lower bounds for Schur and weak Schur numbers

### 2021-10-28 article/article.tex
modifs :O
oups

This article provides new lower bounds for both Schur and weak Schur numbers by exploiting a "template"-based approach. 
The concept of "template" is also generalized to weak Schur numbers. Finding new templates leads to explicit partitions 
improving lower bounds as well as the growth rate for Schur numbers, weak Schur numbers and multicolor Ramsey numbers \(R_n(3)\). 
The new lower bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS(6) \geqslant 646\), 
\(\WS (9) \geqslant 22\,536\) and \(\WS (10) \geqslant 71\,256 \).

>>>>

This article provides new lower bounds for both Schur and weak Schur numbers by exploiting a "template"-based approach. 
The concept of "template" is also generalized to weak Schur numbers. Finding new templates leads to explicit partitions 
improving lower bounds as well as the growth rate for Schur numbers, weak Schur numbers, and multicolor Ramsey numbers \(R_n(3)\). 
The new lower bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS(6) \geqslant 646\), 
\(\WS (9) \geqslant 22\,536\) and \(\WS (10) \geqslant 71\,256 \).

### 2021-10-28 article/schur.tex
oups

Here \(n\) can be seen as a "special" color in the sense that it is no constrained by \ref{eq:constemp}.
However, note that \(n\) is not necessarily the last color by order of appearance, any color can play this role.

@@

\begin{proof}[\textsc{Proof.}]
The upper bound comes from the fact that an S-template with width \(p\) and \(n\) colors is also a partition of \([\![1, p]\!]\) 
into \(n\) sum-free subsets. The lower bound comes from Abbott and Hanson's construction \cite{AbbottHanson}. \\
\end{proof}

@@

We start by reminding the explicit construction of a sum-free partition with the use of an S-template and another sum free partition. 
We rephrase in terms of Schur numbers this contruction stated by Rowley in the context of Ramsey numbers \cite{RowleyRamsey}.

\begin{theorem}
\label{thm:Stemp}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there is a \(S\)-template with width \(q\) and \(n+1\) colors 
	and a sum-free \(k\)-partition of \([\![1,p]\!]\) then there is a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) denotes the minimum element colored with the special color in the \(S\)-template.
\end{theorem}

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They 
extended a sum-free partition vertically by repeating it. Next, they used another sum-free partition to color the other half according 
to the line number, as in \hyperref[figure:1]{Figure 1}. We give an example for \(p = 4\), 
\(q = 9\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) \left(2 S(2) + 1\right) + S(2)\), both with Abbott and 
Hanson's construction (\hyperref[figure:1]{Figure 1}) and with an S-template which is not included in Abbott and Hanson's construction 
(\hyperref[figure:2]{Figure 2}). In both cases, the special color is grey.

@@

Setting \(q = S^+(n+1)\) and \(p = S(k)\) in theorem \ref{thm:Stemp} yields the following corollary.

@@

A S-template with \(n\) colors is said to be symmetric if the sum-free \(n\)-partition derived 
from this template by applying the extension procedure with a sum-free partition of length 1 is symmetric. 
\end{definition}

We produced S-templates using the lingeling SAT solver \cite{Lingeling2017}, hence providing lower bound on 
\(S^+\) and inequalities of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing us with 
the largest value of \((a, b)\) (in the lexicographic order). In order to reduce the search space when the number 
of colors exceeded five we only looked for symmetric S-templates, we assumed that the special color was the last 
color to appear and we constrained the \(m_c\)'s out of being too small. Details concerning the encoding as a SAT 
problem can be found in \cite{Heule2017}.

@@

Inequality (\ref{S(n+1)}) comes from  Schur's original article\cite{Schur1917}. Inequality (\ref{S(n+2)}) is due to
Abott and Hanson \cite{AbbottHanson} and inequality (\ref{S(n+3)}) to Rowley \cite{RowleyRamsey}. 
The three following inequalities are our result.

@@

\par The \hyperref[S-templates]{templates} corresponding to inequalities (\ref{S(n+3)}), (\ref{S(n+4)}) and 
(\ref{S(n+5)}) are listed in \ref{S-templates}.

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}) and (\ref{S(n+3)}) cannot be further improved (with this definition of S-template). 
Inequality (\ref{S(n+4)}) cannot be further improved by only searching for symmetric S-templates whose special color is the 
last in the order of apparition (and with a multiplicative factor less than or equal to 118). Inequality (\ref{S(n+5)}) can most 
likely be further improved but the improvement probably will not be substantial. Finally, a S-template corresponding to inequality 
(\ref{S(n+6)}) was found by extending into a S-template the Schur 6 partition used in \cite{rowley2021improved} (owing to the 
size of this template, its interest is limited and since it can easily be derived from above mentioned partition, it is not given in 
\ref{S-templates}). Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) can definitely be 
improved by a wide margin.

@@

Except for \(S(8)\), \(S(9)\) and \(S(13)\), the best lower bounds are obtained thanks to
the fifth inequality \( S(n+5) \geqslant 380 \, S(n) + 148\). The table
does not go any further, but the same inequality allows one to improve the
lower bounds for \( n \geqslant 15 \).

>>>>

Here \(n\) can be seen as a "special" color in the sense that it is not constrained by Assertion (\ref{eq:constemp}).
However, note that \(n\) is not necessarily the last color by order of appearance, any color can play this role.

@@

\begin{proof}[\textsc{Proof.}]
The upper bound comes from the fact that an S-template with width \(p\) and \(n\) colors is also a sum-free \(n\)-partition 
of \([\![1, p]\!]\). The lower bound comes from Abbott and Hanson's construction \cite{AbbottHanson}. \\
\end{proof}

@@

We start by reminding the explicit construction of a sum-free partition with the use of an S-template and a sum-free partition. 
We rephrase in terms of Schur numbers this contruction stated by Rowley in the context of Ramsey numbers \cite{RowleyRamsey}.

\begin{theorem}
\label{thm:Stemp}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there is an S-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there is a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. Number \(m_{n+1}\) denotes the minimum element colored with the special color in the \(S\)-template.
\end{theorem}

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. First, they
put the integers from \([\![1,S(k)(2S(n)+1)+S(n)]\!]\) in a table with width \(2S(n)+1\) and height \(S(k)+1\) and they 
numbered both the rows and the columns, starting from one upwards. Then, they 
extended a sum-free partition vertically by repeating it. Next, they used another sum-free partition to color the other half according 
to the line number, as in \hyperref[figure:1]{Figure 1}. We give an example for \(p = 4\), 
\(q = 9\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) \left(2 S(2) + 1\right) + S(2)\), both with Abbott and 
Hanson's construction (\hyperref[figure:1]{Figure 1}) and with an S-template which is not included in Abbott and Hanson's construction 
(\hyperref[figure:2]{Figure 2}). In both cases, the special color is grey.

@@

Setting \(q = S^+(n+1)\) and \(p = S(k)\) in Theorem \ref{thm:Stemp} yields the following corollary.

@@

A S-template with \(n\) colors is said to be symmetric if the partition into \(n\) sum-free subsets derived 
from this template by applying the extension procedure of theorem \ref{thm:Stemp} with a sum-free partition of length 1 is symmetric. 
\end{definition}

We produced S-templates using the lingeling SAT solver \cite{Lingeling2017}, hence providing lower bounds on 
\(S^+\) and inequalities of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing us with 
the largest value of \((a, b)\) (in the lexicographic order). In order to reduce the search space when the number 
of colors exceeded five we only looked for symmetric S-templates, we assumed that the special color was the last 
color to appear and we constrained the \(m_c\)'s out of being too small. Details concerning the encoding as a SAT 
problem can be found in \cite{Heule2017}.

@@

Inequality (\ref{S(n+1)}) comes from  Schur's original article \cite{Schur1917}. Inequality (\ref{S(n+2)}) is due to
Abott and Hanson \cite{AbbottHanson} and inequality (\ref{S(n+3)}) to Rowley \cite{RowleyRamsey}. 
The three following inequalities are our result:

@@

\par The \hyperref[S-templates]{templates} corresponding to inequalities (\ref{S(n+3)}), (\ref{S(n+4)}), and 
(\ref{S(n+5)}) are listed in \ref{S-templates}.

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}), and (\ref{S(n+3)}) cannot be further improved (with this definition of S-template). 
Inequality (\ref{S(n+4)}) cannot be further improved by only searching for symmetric S-templates whose special color is the 
last in the order of apparition (and with a multiplicative factor less than or equal to 118). Inequality (\ref{S(n+5)}) can most 
likely be further improved but the improvement probably will not be substantial. Finally, a S-template corresponding to inequality 
(\ref{S(n+6)}) was found by extending into a S-template the Schur 6-partition used in \cite{rowley2021improved} (owing to the 
size of this template, its interest is limited and since it can easily be derived from above mentioned partition, it is not given in 
\ref{S-templates}). Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) can definitely be 
improved by a wide margin.

@@

Except for \(S(8)\), \(S(9)\), and \(S(13)\), the best lower bounds are obtained thanks to
the fifth inequality \( S(n+5) \geqslant 380 \, S(n) + 148\). The table
does not go any further, but the same inequality allows one to improve the
lower bounds for \( n \geqslant 15 \).

### 2021-10-28 article/weakschur.tex
oups

\begin{theorem}
\label{thm:WS-ab}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there are a weakly sum-free \(n\)-partition of \([\![1,q]\!]\)
and a sum-free \(k\)-partition of \([\![1,p]\!]\) then there is a partition of
\({[\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]}\) into \(n+k\) weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

Given that Theorem \ref{thm:WS-ab} will appear as a particular case of a more general theorem after the introduction of
templates for weak Schur numbers, we only outline the demonstration; a \hyperref[PreuveThm]{proof} using templates 
for weak Schur numbers is in Subsection \ref{ConstructionWS}.

Let \((p, k), (q, n) \in (\mathbb{N}^*)^2\) such that there are a weakly sum-free \(n\)-partition of \([\![1,q]\!]\) 
and a sum-free \(k\)-partition of \([\![1,p]\!]\). Let \(a \in \mathbb{N}\) with \(a > q\)
and let us try to build a coloring of \([\![1, ap + q]\!]\) into \(n + k\) weakly sum-free subsets. Let
\(l = a - q - 1\), \(r \in [\![1,q]\!]\) and \(w = a - l - r - 1 = q - r\).

@@

\begin{itemize}
	\item \(\mathcal{T}\) (the "tail"): In the row indexed by 0: numbers from \([\![1,q]\!]\).
	\item \(\mathcal{R}\) (the "rows"): the integers in columns \(-l\) to \(+r\).
	\item \(\mathcal{C}\) (the "columns"): the integers in the last \(w\) columns.
\end{itemize}

@@

This can be seen as an equivalent for weak Schur numbers of Abott and Hanson's construction for Schur numbers 
\cite{AbbottHanson}. This formula also includes the results of Rowley \cite{RowleyWS} as a special case. For \(n>2\), 
the formula from corollary \ref{cor:ineqWS-S} does not give new lower bounds.

@@

\begin{definition}
Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). We define :
\[ \pi_{a,b}:x \longmapsto (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x \mod a).\]
\end{definition}

@@

\begin{sloppypar}
\(\pi\) is the projection on the first line mentioned in the intuitive explanation. The following four propositions are
calculative properties on \(\pi\) reflecting the behaviour of an element's column number in \hyperref[SchemaWS]{Figure 3} and
that we will use later when we introduce WS-templates.
\end{sloppypar}

@@

Function \(\lambda\) maps an element \(x\) to its line number as mentioned in the intuitive explanation.
As we just did with \(\pi\), we prove three more calculative results on both \(\pi\) and \(\lambda\) that are used in
Subsection \ref{ConstructionWS}.

@@

\begin{definition}
Let \( (a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\). Let \((A_1,...,A_n)\) a partition of  \([\![1, a + b]\!]\).
This partition is said to be a \(b\)-weakly-sum-free template (\(b\)-WS-template) with width \(a\) and \(n\) colors when :

\begin{itemize}
\item \(\forall i \in [\![1, n]\!], A_i\) is weakly-sum-free,
\item \(\forall i \in [\![1, n]\!], A_i\backslash [\![1, b]\!]\) is sum-free,
\item For \(A_n\) (the special subset) :
	\[
	\forall (x,y) \in A_n^2, \,x+y>b+2a \implies x+y-2a\notin A_n,
	\]
\item For the others subsets
	\[
	\forall i \in [\![1,n-1]\!], \, \forall(x,y) \in A_i^2, \, x+y>a+b \implies \pi(x+y) \notin A_i.
	\]
\end{itemize}
\end{definition}

@@

\begin{definition}
Let \( (n,b) \in (\mathbb{N}^*)^2\). If there is \(a\) such that there is a \(b\)-WS-template with width \(a\)
and \(n\) colors, we define :
\[
\WS_b^+(n)= \max \{a \in \mathbb{N}^*/ \text{there is a } b \text{-WS-template with width } a \text{ and } n \text{ colors} \}.
\]
If there is no such \(a\), we set \(\WS_b^+(n) = 0\).
\end{definition}

\begin{definition}
Let \( n \in \mathbb{N}^*\). We define :
\[
\WS^+(n)=\max_{b\in \mathbb{N}^*} \WS_b^+(n).
\]
\end{definition}
The following proposition briefly shows how \(\WS^+(n)\) compares to weak Schur numbers.
\begin{proposition}
Let \(n \in [\![2, +\infty]\!]\). Then :
\[
\frac{3}{2} \WS (n-1)+1 \leqslant \WS^+(n) \leqslant \WS (n).
\]
\end{proposition}

@@

\begin{theorem}
\label{thm:WStemp}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there are a sum-free \(k\)-partition of
\([\![1,p]\!]\) and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there is a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

@@

We split \([\![1, p a + b]\!]\) into three subsets. \\
NB: To keep the notation short, the restriction to \([\![b + 1, p a + b]\!]\) of \(\pi\) defined in Subsection \ref{DefinitionWS+}
is denoted by \(\pi\) in the hereunder equations.

\begin{itemize}
	\item \(\mathcal{T} = [\![1, b]\!]\)
	\item \(\mathcal{C} = \pi^{-1}(f^{-1}([\![1, n]\!]))\)
	\item \(\mathcal{R} = \pi^{-1}(f^{-1}(\{n + 1\}))\)
\end{itemize}

We define a coloring \(h\) of \([\![1, p a + b]\!]\) using \( [\![1,n+k]\!]\) colors:
\[
\begin{array}{c c c l}
	h : & [\![1, p a + b]\!] & \longrightarrow & [\![1,n+k]\!] \\
	& x & \longmapsto &
	\left\{ \begin{array}{l l}
		f(x) & \text{if}~x \in \mathcal{T} \\
		f(\pi(x)) & \text{if}~x \in \mathcal{C} \\
		n + g(\lambda(x)) & \text{if}~x \in \mathcal{R}
	\end{array} \right.
\end{array}
\]

@@

Theorem \ref{thm:WS-ab} can be seen as a particular case of Theorem \ref{thm:WStemp} in the same way Abott and Hanson's 
construction \cite{AbbottHanson} can be seen as a particular case of theorem \ref{thm:Stemp}.

@@

\[
\begin{array}{c c c l}
	g : & [\![1, a + b]\!] & \longrightarrow & [\![1, n + 1]\!] \\
	& x & \longmapsto &
		\left\{ \begin{array}{l l}
			f(x) & \text{if}~x \in [\![1, b]\!]  \\
			n + 1 & \text{if}~x \in [\![b + 1, 2 b + 1]\!] \\
			f(x - a) & \text{if}~x \in  [\![2 b + 2, a + b]\!]
		\end{array} \right.
\end{array}
\]

@@

\begin{corollary}
\label{cor:ineqWS+}
Let \(n,k \in \mathbb{N}^*\) and set \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then :
\[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}.
\]
\end{corollary}

@@

\par
These modifications rely on one main observation: the first row (excluding the "tail") has constraints that other rows do not 
have because of the tail, especially if the special color appears in the tail as well. Therefore, having a coloring dedicated to 
the first row and another coloring dedicated to the repeating pattern would weaken the constraints on the repeating pattern.

@@

We produced WS-templates using the lingeling SAT solver \cite{Lingeling2017}, hence providing lower bound on \(WS^+\) 
and inequalities of the type \(WS(n+k) \geqslant a S(n) + b\). We sought templates providing the greatest value of \((a, b)\) 
(in the lexicographic order). Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

@@

Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates with five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions.

Like in  Subsection \ref{subsec:lowS}, we compute the lower bounds given by inequalities (\ref{WS(n+1)}),
(\ref{WS(n+2)}) and (\ref{WS(n+3)}) for \( n \in [\![8,15]\!] \). The best lower bound for each value of \(n\) is highlighted.

>>>>

\begin{theorem}
\label{thm:WS-ab}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there are a weakly sum-free \(n\)-partition of \([\![1,q]\!]\) and
a sum-frree \(k\)-partition of \([\![1,p]\!]\) then there is a partition of \({[\![1,p(q+\left \lceil \frac{q}{2} \right 
\rceil + 1)+q]\!]}\) into \(n+k\) weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

As Theorem \ref{thm:WS-ab} is actually a particular case of a more general theorem (Theorem \ref{thm:WStemp}) which will
be formulated after the introduction of templates for weak Schur number,
we only outline the demonstration. Its complete \hyperref[PreuveThm]{proof} using templates 
for weak Schur numbers is in Subsection \ref{ConstructionWS}.

Let \((p, k), (q, n) \in (\mathbb{N}^*)^2\) such that there are a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets. Let \(a \in \mathbb{N}\) with \(a > q\)
and let us try to build a coloring of \([\![1, ap + q]\!]\) into \(n + k\) weakly sum-free subsets. Let
\(l = a - q - 1\), \(r \in [\![1,q]\!]\) and \(w = a - l - r - 1 = q - r\).

@@

\begin{itemize}
	\item \(\mathcal{T}\) (the "tail"): in the row indexed by 0: numbers from \([\![1,q]\!]\).
	\item \(\mathcal{R}\) (the "rows"): the integers in columns from \(-l\) to \(+r\).
	\item \(\mathcal{C}\) (the "columns"): the integers in the last \(w\) columns.
\end{itemize}

@@

This can be seen as an equivalent for weak Schur numbers of Abott and Hanson's construction for Schur numbers 
\cite{AbbottHanson}. This formula also includes the results from \cite{RowleyWS} as a special case. For \(n>2\), 
the formula from Corollary \ref{cor:ineqWS-S} does not give new lower bounds.

@@

\begin{definition}
Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). We define:
\[ \pi_{a,b}:x \longmapsto (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x \mod a).\]
\end{definition}

@@

\begin{sloppypar}
Function \(\pi\) is the projection on the first line mentioned in the intuitive explanation. The following four propositions are
calculative properties on \(\pi\) reflecting the behaviour of an element's column number in \hyperref[SchemaWS]{Figure 3} and
that we will use later when we introduce WS-templates.
\end{sloppypar}

@@

Function \(\lambda\) maps an element \(x\) to its line number as mentioned in the intuitive explanation.
As we have just done with \(\pi\), we prove three more calculative results on both \(\pi\) and \(\lambda\) that are used in
Subsection \ref{ConstructionWS}.

@@

\begin{definition}
Let \( (a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\). Let \((A_1,...,A_n)\) a partition of  \([\![1, a + b]\!]\).
This partition is said to be a \(b\)-weakly-sum-free template (\(b\)-WS-template) with width \(a\) and \(n\) colors when:

\begin{itemize}
\item \(\forall i \in [\![1, n]\!], A_i\) is weakly-sum-free,
\item \(\forall i \in [\![1, n]\!], A_i\backslash [\![1, b]\!]\) is sum-free,
\item For \(A_n\) (the special subset):
	\[
	\forall (x,y) \in A_n^2, \,x+y>b+2a \implies x+y-2a\notin A_n,
	\]
\item For the others subsets
	\[
	\forall i \in [\![1,n-1]\!], \, \forall(x,y) \in A_i^2, \, x+y>a+b \implies \pi(x+y) \notin A_i.
	\]
\end{itemize}
\end{definition}

@@

\begin{definition}
Let \( (n,b) \in (\mathbb{N}^*)^2\). If there is \(a\) such that there is a \(b\)-WS-template with width \(a\)
and \(n\) colors, we define:
\[
\WS_b^+(n)= \max \{a \in \mathbb{N}^*/ \text{there is a } b \text{-WS-template with width } a \text{ and } n \text{ colors} \}.
\]
If there is no such \(a\), we set \(\WS_b^+(n) = 0\).
\end{definition}

\begin{definition}
Let \( n \in \mathbb{N}^*\). We define:
\[
\WS^+(n)=\max_{b\in \mathbb{N}^*} \WS_b^+(n).
\]
\end{definition}
The following proposition briefly shows how \(\WS^+(n)\) compares to weak Schur numbers.
\begin{proposition}
Let \(n \in [\![2, +\infty]\!]\). Then:
\[
\frac{3}{2} \WS (n-1)+1 \leqslant \WS^+(n) \leqslant \WS (n).
\]
\end{proposition}

@@

\begin{theorem}
\label{thm:WStemp}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there are a partition of \([\![1,p]\!]\)
into \(k\) sum-free subsets and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there is a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

@@

We split \([\![1, p a + b]\!]\) into three subsets. \\
NB: To keep the notation short, the restriction to \([\![b + 1, p a + b]\!]\) of \(\pi\) defined in Subsection \ref{DefinitionWS+}
is denoted by \(\pi\) in the hereunder equations:

\begin{itemize}
	\item \(\mathcal{T} = [\![1, b]\!]\),
	\item \(\mathcal{C} = \pi^{-1}(f^{-1}([\![1, n]\!]))\),
	\item \(\mathcal{R} = \pi^{-1}(f^{-1}(\{n + 1\}))\).
\end{itemize}

We define a coloring \(h\) of \([\![1, p a + b]\!]\) using \( [\![1,n+k]\!]\) colors:
\[
\begin{array}{c c c l}
	h : & [\![1, p a + b]\!] & \longrightarrow & [\![1,n+k]\!],\\
	& x & \longmapsto &
	\left\{ \begin{array}{l l}
		f(x), & \text{if}~x \in \mathcal{T}, \\
		f(\pi(x)), & \text{if}~x \in \mathcal{C}, \\
		n + g(\lambda(x)), & \text{if}~x \in \mathcal{R}.
	\end{array} \right.
\end{array}
\]

@@

We go back to Theorem \ref{thm:WS-ab} which can be seen as a particular case of Theorem \ref{thm:WStemp} in the same way Abott and Hanson's 
construction \cite{AbbottHanson} can be seen as a particular case of Theorem \ref{thm:Stemp}. Its deferred proof is given below.

@@

\[
\begin{array}{c c c l}
	g : & [\![1, a + b]\!] & \longrightarrow & [\![1, n + 1]\!], \\
	& x & \longmapsto &
		\left\{ \begin{array}{l l}
			f(x), & \text{if}~x \in [\![1, b]\!],  \\
			n + 1, & \text{if}~x \in [\![b + 1, 2 b + 1]\!], \\
			f(x - a), & \text{if}~x \in  [\![2 b + 2, a + b]\!].
		\end{array} \right.
\end{array}
\]

@@

\begin{corollary}
\label{cor:ineqWS+}
Let \(n,k \in \mathbb{N}^*\) and set \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then:
\[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}.
\]
\end{corollary}

@@

\par
These modifications rely on one main observation: the first row (excluding the "tail") has constraints that other rows do not 
have because of the tail, especially if the special color appears in the tail as well. Indeed, the line number of the
sum of two numbers in the tail is never greater than one. Therefore, having a coloring dedicated to 
the first row and another coloring dedicated to the repeating pattern would weaken the constraints on the repeating pattern.

@@

We produced WS-templates using the lingeling SAT solver \cite{Lingeling2017}, hence providing lower bounds on \(WS^+\) 
and inequalities of the type \(WS(n+k) \geqslant a S(n) + b\). We sought templates providing the greatest value of \((a, b)\) 
(in the lexicographic order). Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

@@

Inequality (\ref{WS(n+3)}), found with a SAT solver \cite{Lingeling2017}, cannot be improved (with this 
definition of WS-template). It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates with five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We expect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions.

Like in  Subsection \ref{subsec:lowS}, we compute the lower bounds given by Inequalities (\ref{WS(n+1)}),
(\ref{WS(n+2)}), and (\ref{WS(n+3)}) for \( n \in [\![8,15]\!] \). The best lower bound for each value of \(n\) is highlighted.
